---
title: "pink-salmon-forecast-notebook"
author: "Dan Ovando"
date: "3/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## launch call 2020-03-05

Where do they see major areas for improvement?

DEscribe the validation scheme in those forecast plots

How much data realistically are we looking at here?

Reason forecasting harvest not numbers?

Andy piston

Another dataset of chum salmon to discuss

weak links - that survey takes place in one place and trying some experimental new trawls in lower chatahm and sumner straight. 

Once those fish go out to sea information on juvenile abundance in southern southeast alaska

Watson didn't find much that panned out, but did try satellite data, but just from that area?


JUvenile survey did an august survey a june july and a may, but almost never in may

but the peak is in june or july 

missing the peak might be an issue with chum salmon

survey started in 98

reliable forecasts by subregion would be valuable 

1960


no offical forecast for chum (90% hatchery fish), and the hatcheris all do their own forecasts

blob years mammoth returns of 3's, which messed up the forecast since things just came back early (bit 3 did not equal big 4)

Can get us in contact with the Chum world 

chum value has gone up a lot for processors - a lot more targeted fisheries for chum, particularly in bad pink years

Data & survey are readily available NPASC documents

Chums most of the information you'd have to go through the hatchery operators themselves, but they would probably be interested in collaborating 
  
DAIPAC

marine envirnomental stuff moight have explained high survival in the boom years


# exploring RAM


```{r}
  
library(here)
library(tidyverse)
 if (file.exists(here("data", "ram.zip")) == FALSE) {
    download.file(
      "https://zenodo.org/record/4824192/files/RAMLDB%20v4.495.zip?download=1",
      destfile = here("data", "ram.zip"),
      mode = "wb"
    )

    unzip(here("data", "ram.zip"), exdir = here("data", "ram")) # unzip = 'unzip' needed for windows
  }

  ram_files <-
    list.files(here("data", "ram", "R Data"), recursive = TRUE)

  ram_files <- ram_files[str_detect(ram_files, ".RData")]

  load(here("data", "ram", "R Data", ram_files[1]))

  # process ram data ------------------------------------------------------------

  stock <- stock %>%
    left_join(area, by = "areaid")
  
  pink_stocks <- stock %>% 
    filter(str_detect(tolower(scientificname),"oncorhynchus") & str_detect(tolower(commonname), "pink"))

  # catches
  ram_catches <- tcbest.data %>%
    mutate(year = rownames(.) %>% as.integer()) %>%
    as_tibble() %>%
    gather(stockid, catch, -year)

  # B/Bmsy
  ram_b_v_bmsy <- divbpref.data %>%
    mutate(year = rownames(.) %>% as.integer()) %>%
    tibble() %>%
    gather(stockid, b_v_bmsy, -year)


  # U/Umsy
  ram_u_v_umsy <- divupref.data %>%
    mutate(year = rownames(.) %>% as.integer()) %>%
    as_tibble() %>%
    gather(stockid, u_v_umsy, -year)

  # Effort
  ram_effort <- effort.data %>%
    mutate(year = rownames(.) %>% as.integer()) %>%
    as_tibble() %>%
    gather(stockid, effort, -year)

  # biomass

  ram_total_biomass <- tbbest.data %>%
    mutate(year = rownames(.) %>% as.integer()) %>%
    as_tibble() %>%
    gather(stockid, total_biomass, -year)

  
  pink_bio <- ram_total_biomass %>% 
    filter(stockid %in% unique(pink_stocks$stockid))
  # ssb

  ram_ss_biomass <- ssb.data %>%
    mutate(year = rownames(.) %>% as.integer()) %>%
    as_tibble() %>%
    gather(stockid, ss_biomass, -year)


   pink_ssb <- ram_ss_biomass %>% 
    filter(stockid %in% unique(pink_stocks$stockid))
   
   a = pink_ssb %>%
     group_by(stockid) %>%
     mutate(has = any(!is.na(ss_biomass)),
            ssb = scale(ss_biomass),
            even = ifelse((year %% 2) == 0, "Even", "Odd")) %>%
     ungroup() %>%
     filter(has,
            year > 1960) %>%
     ggplot(aes(year, ssb, color = stockid)) +
     geom_line(show.legend = FALSE, alpha = 0.5) + 
     facet_wrap(~even)
   
   pink_catches <- ram_catches %>% 
         filter(stockid %in% unique(pink_stocks$stockid))

  # ssb
  
  ram_exp_rate <- ram_catches %>%
    left_join(ram_total_biomass, by = c("stockid", "year")) %>%
    mutate(exploitation_rate = catch / total_biomass) %>%
    select(-catch,-total_biomass)

  # put it together

  ram_data <- ram_catches %>%
    left_join(bioparams_values_views, by = "stockid") %>%
    left_join(ram_b_v_bmsy, by = c("stockid", "year")) %>%
    left_join(ram_u_v_umsy, by = c("stockid", "year")) %>%
    left_join(ram_exp_rate, by = c("stockid", "year")) %>%
    left_join(ram_effort, by = c("stockid", "year")) %>%
    left_join(ram_total_biomass, by = c("stockid", "year")) %>%
    left_join(ram_ss_biomass, by = c("stockid", "year")) %>%
    left_join(stock, by = "stockid") %>%
    select(stockid, scientificname, commonname, everything())


  # create new variables

  ram_data <- ram_data %>%
    mutate(tb_v_tb0 = total_biomass / TB0,
           ssb_v_ssb0 = ss_biomass / SSB0)

  


```

# Check in with Jenny Morella (prince william sound)


general overvie of pinks in PWS

two different hatechery organizations, pWS unique in that hatchery is a big role in pinks

Individual hatchery origin is findable

Otololith lab is used to separate out hatchery vs. wild and which hatchery for the hatchery part

If it's unmarked, it's wild. 

Harvest component comes from that. Aerial survey informs the escapement. 

Pink salmon stream length, observer efficiency per stream, and is used as the basis of an aerial survey program. Fly index streams, and then expand based on streams not observed and efficiency 

In 2015 survey program was cut so reduced aerial survey down to a smaller number, messing with the old expansion thingy. 

Escapement and escapement goals - 

So the important thing is that there's a change in 2015 that makes things not directly comparable

1960 is the total run

Annual try relesase and hatchery forecasts are available on the hatchery group websites. Assume a fixed marine survival and basically extrapolate from fry release. 

"a clearly defined geographical space, recognised, dedicated and managed, through legal or other effective means, to achieve the long-term conservation of nature with associated ecosystem services and cultural values"


# Pinks

Looking at precipitation as an effect 

even or odd as two different models

warm and wet years are good for pinks


